---
title: "Chapter_7_polynoimial_step_piecewise_polynomial"
author: "Ruijuan Li"
date: "3/21/2018"
output: 
  html_document: 
    keep_md: yes
---

### 6 

6. In this exercise, you will further analyze the Wage data set considered throughout this chapter.

(a) Perform polynomial regression to predict wage using age. Use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the results of hypothesis testing using ANOVA? Make a plot of the resulting polynomial fit to the data.

```{r}
library(tidyverse)
library(ISLR)
library(gam) 

# fit model with different degrees, and use anova 
fit.1=lm(wage~age,data=Wage)
fit.2=lm(wage~poly(age,2),data=Wage) 
fit.3=lm(wage~poly(age,3),data=Wage) 
fit.4=lm(wage~poly(age,4),data=Wage) 
fit.5=lm(wage~poly(age,5),data=Wage) 
anova(fit.1,fit.2,fit.3,fit.4,fit.5) # degree three I prefer 

# use cross validation 
library(boot)
set.seed(1)
cv.error.10=rep(0,5) 
for (i in 1:5){
glm.fit=glm(wage~poly(age ,i),data=Wage) 
cv.error.10[i]=cv.glm(Wage,glm.fit,K=10)$delta[1] # 10 fold CV 
} 
cv.error.10 # degree four is the best 

# make a plot for polynomial, I prefer result from cv 
# get predict 
agelims=range(Wage$age)
age.grid=seq(from=agelims[1],to=agelims[2])
preds=predict(fit.3,newdata=list(age=age.grid),se=TRUE) # predict
se.bands=cbind(preds$fit+2*preds$se.fit,preds$fit-2*preds$se.fit ) # predict +/- 2se

# plot 
plot(Wage$age,Wage$wage,xlim=agelims ,cex=.5,col="darkgrey") # plot the value
title("Degree -4 Polynomial ",outer=T) # title 
lines(age.grid,preds$fit,lwd=2,col="blue") # add fit line 
matlines(age.grid,se.bands,lwd=1,col="blue",lty=3) # add CI line 
```

(b) Fit a step function to predict wage using age, and perform cross- validation to choose the optimal number of cuts. Make a plot of the fit obtained.

```{r}
# cv, for each cut from 2 to 6, run 10 times validation set test, each validation set start with a different seed 
# make a function 
deltas <- rep(NA, 9)
for (i in 2:10) {
    Wage$age.cut <- cut(Wage$age, i) 
    fit <- glm(wage ~ age.cut, data = Wage)
    deltas[i-1] <- cv.glm(Wage, fit, K = 10)$delta[1] ### 
    names(deltas)[i-1] <- paste0("Cuts = ", i-1)
}
deltas
plot(1:9, deltas, xlab = "Cuts", ylab = "Test MSE", type = "l")
(d.min <- which.min(deltas))
points(d.min, deltas[d.min], col = "red", cex = 2, pch = 20)   

# check this  
```

### 7 

The Wage data set contains a number of other features not explored in this chapter, such as marital status (maritl), job class (jobclass), and others. Explore the relationships between some of these other predictors and wage, and use non-linear fitting techniques in order to fit flexible models to the data. Create plots of the results obtained, and write a summary of your findings.

```{r, error=TRUE}
library(ISLR) 
colnames(Wage)  
pairs(Wage)

Wage$maritl <- as.numeric(Wage$maritl)  
# use maritial status to predict wage, categorical predictor, I decided to use step function  
table(cut(Wage$maritl,4)) # cut pick cutpoint automatically or specify cut point using breaks 
fit=lm(wage~cut(maritl ,4),data=Wage) 
coef(summary(fit))

# predict 
agelims=range(Wage$age) 
age.grid=seq(from=agelims[1],to=agelims[2])
preds=predict(fit,newdata=list(maritl=age.grid),se=TRUE) # predict
se.bands=cbind(preds$fit+2*preds$se.fit,preds$fit-2*preds$se.fit ) # predict +/- 2se

# plot the fit result 
plot(Wage$maritl,Wage$wage,xlim=agelims ,cex=.5,col="darkgrey") # plot the value
title("Step fit marital ",outer=T) # title 
lines(age.grid,preds$fit,lwd=2,col="blue") # add fit line 
matlines(age.grid,se.bands,lwd=1,col="blue",lty=3) # add CI line  

### update with smooth splines & GAM 
# smooth spline, not be able to fit on factors predictors 
agelims=range(Wage$year) 
age.grid=seq(from=agelims[1],to=agelims[2])
plot(Wage$year,Wage$wage,xlim=agelims ,cex=.5,col="darkgrey")
title (" Smoothing Spline ")
fit=smooth.spline(Wage$year,Wage$wage,df=6)
fit2=smooth.spline(Wage$year,Wage$wage,cv=TRUE)
fit2$df
lines(fit,col="red",lwd=2) # smooth spline with df of 16 
lines(fit2,col="blue",lwd=2)
legend("topright",legend=c("6 DF","2 DF"), 
col=c("red","blue"),lty=1,lwd=2,cex=.8) # add legend 

# if fit on marital status 
fit=smooth.spline(Wage$maritl,Wage$wage,df=16)

# local regression, loess(), not be able to fit on maritl or education...  
plot(Wage$year,Wage$wage,xlim=agelims ,cex=.5,col="darkgrey")
title (" Local Regression ")
fit=loess(wage~year,span=.5,data=Wage)
fit2=loess(wage~year,span=.8,data=Wage)
lines(age.grid,predict(fit,data.frame(year=age.grid)), col="red",lwd=2)
lines(age.grid,predict(fit2,data.frame(year=age.grid)), col="blue",lwd=2)
legend("topright",legend=c("Span=0.5","Span=0.8"), col=c("red","blue"),lty=1,lwd=2,cex=.8) 

# GAM, can replace  
gam.m1=gam(wage~s(age ,5)+maritl ,data=Wage) # model that exclude year 
gam.m2=gam(wage~year+s(age ,5)+maritl ,data=Wage) # model that use linear function 
gam.m3=gam(wage~s(year ,4)+s(age ,5)+maritl ,data=Wage)

anova(gam.m1,gam.m2,gam.m3,test="F") 

plot.gam(gam.m2, se=TRUE,col="blue")  
```

### 8 

Fit some of the non-linear models investigated in this chapter to the Auto data set. Is there evidence for non-linear relationships in this data set? Create some informative plots to justify your answer.

```{r}
colnames(Auto) 

# I would like to invesitigate the relationship between origin and mpg, hypothesis, Japanese car has a higher mpg, origin matters 

# polynomial 
fit=lm(mpg~poly(origin,2),data=Auto) 
summary(fit) # degree of two 

# step 
table(cut(Auto$origin, breaks = 4)) # cut pick cutpoint automatically or specify cut point using breaks 
fit=lm(mpg~cut(origin ,4),data=Auto)  
coef(summary(fit))

# predict 
agelims=range(Auto$origin)
age.grid=seq(from=agelims[1],to=agelims[2])
preds=predict(fit,newdata=list(origin=age.grid),se=TRUE) # predict
se.bands=cbind(preds$fit+2*preds$se.fit,preds$fit-2*preds$se.fit ) # predict +/- 2se

# plot the fit result 
plot(Auto$origin,Auto$mpg,xlim=agelims ,cex=.5,col="darkgrey") # plot the value
title("Step fit mpg vs. origin ",outer=T) # title 
lines(age.grid,preds$fit,lwd=2,col="blue") # add fit line 
matlines(age.grid,se.bands,lwd=1,col="blue",lty=3) # add CI line, origin matters 

# splines 
library(splines)
fit=lm(mpg~bs(origin,knots=c(1,2,3)),data=Auto) # specify knots at 25, 40, and 60 
pred.splines=predict(fit,newdata=list(origin=age.grid),se=T)
pred.splines 

plot(Auto$origin,Auto$mpg,col="gray")
lines(age.grid,pred.splines$fit,lwd=2)
lines(age.grid,pred.splines$fit+2*pred.splines$se ,lty="dashed")
lines(age.grid,pred.splines$fit-2*pred.splines$se ,lty="dashed")

# natural splines
fit2=lm(mpg~ns(origin,df=4),data=Auto)
pred2=predict(fit2,newdata=list(origin=age.grid),se=T) 
lines(age.grid, pred2$fit,col="red",lwd=2)  

# smooth spline, local regression, gam, interested in weight, year, and origin's relationship with mpg  
pairs(Auto)

agelims=range(Auto$weight) 
age.grid=seq(from=agelims[1],to=agelims[2])
plot(Auto$weight,Auto$mpg,xlim=agelims ,cex=.5,col="darkgrey")
title (" Smoothing Spline ")
fit=smooth.spline(Auto$weight,Auto$mpg,df=6)
fit2=smooth.spline(Auto$weight,Auto$mpg,cv=TRUE)
fit2$df
lines(fit,col="red",lwd=2) # smooth spline with df of 16 
lines(fit2,col="blue",lwd=2)
legend("topright",legend=c("6 DF","10 DF"), 
col=c("red","blue"),lty=1,lwd=2,cex=.8) # add legend 

# local regression, loess(), not be able to fit on maritl or education...  
plot(Auto$weight,Auto$mpg,xlim=agelims ,cex=.5,col="darkgrey")
title (" Local Regression ")
fit=loess(mpg~weight,span=.2,data=Auto)
fit2=loess(mpg~weight,span=.5,data=Auto)
lines(age.grid,predict(fit,data.frame(weight=age.grid)), col="red",lwd=2)
lines(age.grid,predict(fit2,data.frame(weight=age.grid)), col="blue",lwd=2)
legend("topright",legend=c("Span=0.2","Span=0.5"), col=c("red","blue"),lty=1,lwd=2,cex=.8) 

# GAM, can replace
gam.m3=gam(mpg~s(weight ,4)+s(year ,5)+origin ,data=Auto)
par(mfrow=c(1,3))
plot(gam.m3, se=TRUE,col="blue") # except for year, all linear 

gam.m2=gam(mpg~weight+s(year ,5)+origin ,data=Auto) # model that use linear function 

AIC(gam.m2, gam.m3)
BIC(gam.m2, gam.m3)

plot.gam(gam.m3, se=TRUE,col="blue")   
```

### 9 

This question uses the variables dis (the weighted mean of distances to five Boston employment centers) and nox (nitrogen oxides concen- tration in parts per 10 million) from the Boston data. We will treat dis as the predictor and nox as the response.

(a) Use the poly() function to fit a cubic polynomial regression to predict nox using dis. Report the regression output, and plot the resulting data and polynomial fits.

```{r}
library(MASS)
colnames(Boston)

fit=lm(nox~poly(dis,3),data=Boston) 
summary(fit) 

lims=range(Boston$dis)
grid=seq(from=lims[1],to=lims[2])
preds=predict(fit,newdata=list(dis=grid),se=TRUE) # predict
se.bands=cbind(preds$fit+2*preds$se.fit,preds$fit-2*preds$se.fit ) # predict +/- 2se

plot(Boston$dis,Boston$nox,xlim=lims ,cex=.5,col="darkgrey") # plot the value
title("Degree -3 Polynomial ",outer=T) # title 
lines(grid,preds$fit,lwd=2,col="blue") # add fit line 
matlines(grid,se.bands,lwd=1,col="blue",lty=3) # add CI line  
```

(b) Plot the polynomial fits for a range of different polynomial degrees (say, from 1 to 10), and report the associated residual sum of squares.

```{r}
# what is rss? residual: the difference between the ith observed response value and the ith response value that is predicted by our linear model. RSS is sum of squares of residuals 
set.seed(1)
cv.error.10=rep(0,10) 
for (i in 1:10){
glm.fit=glm(nox~poly(dis ,i),data=Boston) 
cv.error.10[i]= sum((Boston$nox-predict(glm.fit, Boston))^2) 
} 
cv.error.10  
plot(cv.error.10) # degree of 3 
```

(c) Perform cross-validation or another approach to select the opti- mal degree for the polynomial, and explain your results.

```{r}
cv.error=rep(0,5) 
for (i in 1:5){
glm.fit=glm(nox~poly(dis ,i),data=Boston) # different polynomial levels 
cv.error[i]=cv.glm(Boston,glm.fit, K=10)$delta[1]}  
cv.error  
plot(cv.error) # degree of 3 
```

I will stop here. 

(d) Use the bs() function to fit a regression spline to predict nox using dis. Report the output for the fit using four degrees of freedom. How did you choose the knots? Plot the resulting fit.

```{r}
```

(e) Now fit a regression spline for a range of degrees of freedom, and plot the resulting fits and report the resulting RSS. Describe the results obtained.

(f) Perform cross-validation or another approach in order to select the best degrees of freedom for a regression spline on this data. Describe your results.

### 10 

10. This question relates to the College data set.

(a) Split the data into a training set and a test set. Using out-of-state tuition as the response and the other variables as the predictors, perform forward stepwise selection on the training set in order to identify a satisfactory model that uses just a subset of the predictors.

```{r}
set.seed(1)
train_ID <- sample(rownames(College), size = round(nrow(College) * 0.8), replace = F) 
train <- College[rownames(College) %in% train_ID,]
test <- College[!(rownames(College) %in% train_ID),]

colnames(College)
library(leaps)
regfit.fwd=regsubsets(Outstate~.,data=College,nvmax=18, method ="forward")
summary(regfit.fwd)
par(mfrow=c(1,1))
plot(regfit.fwd,scale="bic") # I choosethe model including private, P.undergrad, PhD, perc.alumni, and Expend 
```

(b) Fit a GAM on the training data, using out-of-state tuition as the response and the features selected in the previous step as the predictors. Plot the results, and explain your findings.

```{r}
summary(College) 
gam=gam(Outstate~s(P.Undergrad ,4)+s(PhD ,5)+s(perc.alumni ,5) + s(Expend ,5) + Private ,data=train) # should be able to choose the best parameter for s()
par(mfrow=c(2,3))
plot(gam, se=TRUE,col="blue") # except for year, all linear 

```

(c) Evaluate the model obtained on the test set, and explain the results obtained.

```{r}
preds=predict(gam,newdata=test)
# MSE of test error 
mean((test$Outstate-preds)^2)  
```

(d) For which variables, if any, is there evidence of a non-linear relationship with the response?

```{r}
# all of them? should use cross validation to select the best parameter for s() 
summary(gam) # PhD & Expend 
```

