---
title: "Assignment_01_23_2018"
author: "Ruijuan Li"
date: "1/23/2018"
output: 
  html_document: 
    keep_md: yes
---
* I don't know what I am doing here... 

10. This question should be answered using the Weekly data set, which is part of the ISLR package. This data is similar in nature to the Smarket data from this chapter’s lab, except that it contains 1,089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.

(e) Repeat (d) using LDA. (d, fit the logistic regression model using a training data period from 1990 to 2008, with Lag2 as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 and 2010).)

```{r}
library(MASS)
library(tidyverse)
library(ISLR)

Weekly %>% summary() 
train = (Weekly$Year >= 1990 & Weekly$Year <=  2008)
lda.fit.1=lda(Direction ~ Lag2,data=Weekly, subset=train)  
lda.fit.1

Weekly.2005= Weekly[!train,]
lda.pred.1=predict(lda.fit.1, Weekly.2005) 

Direction.2005=Weekly$Direction[!train]
lda.class.1=lda.pred.1$class
table(lda.class.1 ,Direction.2005)

mean(lda.class.1==Direction.2005)
sum(lda.pred.1$posterior[,1]>=.5) 
sum(lda.pred.1$posterior[,1]<.5)

lda.pred.1$posterior[1:20,1] 
lda.class.1[1:20]

sum(lda.pred.1$posterior[,1]>.9)   
```

11. In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the Auto data set.
 
(d) Perform LDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained? 

```{r}
# using weight, displacement, horsepower, ang cylinders as the predictors 

Auto %>% summary()

# (a) Create a binary variable, mpg01, that contains a 1 if mpg contains a value above its median, and a 0 if mpg contains a value below its median. You can compute the median using the median() function. Note you may find it helpful to use the data.frame() function to create a single data set containing both mpg01 and the other Auto variables.

attach(Auto)
mpg01 <- rep(0, length(mpg))
mpg01[mpg > median(mpg)] <- 1
Auto2 <- data.frame(Auto, mpg01)

lda.fit.2=lda(mpg01 ~ weight + displacement + horsepower + cylinders,data=Auto2)  
lda.fit.2 
```

